---
title: "Week 1"
author: "Deo Rupali Vilas"
date: "2026-05-18"
output:
  word_document: default
  html_document: default
---
#Task 1 Data Collection & Dataset Understand
```{r}
data=read.csv("C:/M.Sc/Sample - Superstore.csv")
data
str(data)

summary(data)

dim(data)
names(data)
```
```{r}
unique(data$Ship.Mode)
table(data$Ship.Mode)

unique(data$Customer.Name)
table(data$Customer.Name)

unique(data$Sub.Category)
table(data$Sub.Category)

unique(data$Category)
table(data$Category)
```

```{r}
unique(data$State)
table(data$State)

unique(data$State)
table(data$State)
```

#Discribe:

This dataset represent real-word retail store sales data. in contains information about customer order, product, sales, profit, category, region, and order dates. the dataset helps in customer behavior and sales trend, business performance.

#Data understanding:

This dataset collected are kaggle.This data are 9994 observation and 21 Variable. 15 coloumn are Char and 3 Coloumn ahe integer and 3 coloumn are numeric. this data set are 17 are sub.Category and 3 category. Understanding the structurenof the dataset, identifying column , data types, and recognizing what each feature represents.

#Exploratory thinking:
Exploring the dataset ti identify patterns, trends,relationship,and useful insights from the data.


#Task 2 Data Cleaning & Preprocessing

```{r}
colSums(is.na(data))
```

```{r}
data[!complete.cases(data), ]
data_clean <- na.omit(data)
```


```{r}
sum(duplicated(data))
data <- unique(data)
str(data)
```
#Report Writing
#Missing Values

Missing values in the Superstore dataset were identified using is.na() and removed using na.omit() to improve the quality of the data.

#Duplicate Removal

Duplicate rows were checked using duplicated() and removed using unique() to avoid repeated records.

#Data Formatting

Date columns such as Order Date and Ship Date were converted into proper date format for accurate analysis.


#Task 3: Exploratory Data Analysis (EDA)
```{r}
summary(data)

```

```{r}
mean(data$Sales)
mean(data$Profit)
max(data$Sales)
max(data$Profit)
sd(data$Sales)
sd(data$Profit)
```

```{r}
aggregate(Sales~Category,data=data,mean)
```

```{r}
aggregate(Profit ~ Region, data=data, sum)
```

```{r}
aggregate(Sales ~ Category, data=data, sum)
aggregate(Profit ~ Region, data=data,sum)

```

```{r}
aggregate(Sales ~ Sub.Category, data=data,sum)
```

```{r}
aggregate(Profit ~ Segment, data=data, sum)
```

```{r}
boxplot(data$Sales)
```

```{r}
boxplot(data$Profit)
```

```{r}
boxplot(Profit ~ City, data = data)
```

```{r}
#boxplot.stats(data$Sales)$out
```
#Summarize Findings

The Superstore dataset was analyzed to identify sales and profit patterns across different categories, regions, and cities. The analysis revealed that some categories and cities generated higher profit, while a few showed losses. Basic statistical measures such as mean, minimum, and maximum values helped in understanding the distribution of sales and profit. Boxplots identified several positive and negative outliers, indicating unusually high profits and heavy losses in certain transactions. Overall, the dataset provided useful insights into business performance and sales trends.

#Key Skills
Statistical Thinking Applying statistical methods to understand data distribution, trends, and variations.

#Data Analysis
Analyzing and interpreting business data to discover meaningful insights and patterns.

#Task 4: Data Visualization
```{r}
barplot(tapply(data$Sales, data$Category, sum),
        main="Category-wise Sales",
        col="skyblue")
```

```{r}
barplot(tapply(data$Profit, data$Region, sum),
        main="Region-wise Profit",
        col="lightgreen")
```

```{r}
sales_data <- tapply(df$Sales, df$Category, sum)

bp <- barplot(sales_data,
              col = "skyblue",
              main = "Category-wise Sales")

lines(bp, sales_data,
      type = "b",
      pch = 19,
      col = "red")
```

```{r}
hist(data$Sales,
     main="Sales Distribution",
     xlab="Sales")
```
```{r}
boxplot(data$Profit,
        main="Profit Boxplot")

```
```{r}
# Load Required Libraries
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)

# View Dataset
head(data)

# 1. Sales by Category

sales_category <- data %>%
  group_by(Category) %>%
  summarise(Total_Sales = sum(Sales))

ggplot(sales_category, aes(x = Category, y = Total_Sales, fill = Category)) +
  geom_bar(stat = "identity") +
  ggtitle("Sales by Category") +
  xlab("Category") +
  ylab("Total Sales")

```

# 2. Profit by Region

```{r}
profit_region <- data %>%
  group_by(Region) %>%
  summarise(Total_Profit = sum(Profit))

ggplot(profit_region, aes(x = Region, y = Total_Profit, fill = Region)) +
  geom_bar(stat = "identity") +
  ggtitle("Profit by Region") +
  xlab("Region") +
  ylab("Total Profit")

```

# 3. Monthly Sales Trend

# Key Insights

# Key Skills Used
• Data Visualization  
Created charts and graphs to represent sales, profit, and trends clearly.

• Storytelling with Data  
Analyzed the visualizations and identified important business insights from the dataset.

• Data Analysis  
Performed data exploration and summarized important information from the Superstore dataset.

• Graph Interpretation  
Understood patterns, trends, and regional performance using visual analysis.

#Task 5

```{r}
# Build a simple predictive model using regression analysis

model_data <- data %>%
  select(Sales, Profit)
model_data

# Create Regression Model Predict Profit based on Sales
model <- lm(Profit ~ Sales, data = model_data)
model

# Model Summary
summary(model)

# -----------------------------------
# Visualization
# -----------------------------------

ggplot(model_data, aes(x = Sales, y = Profit)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Sales vs Profit Prediction") +
  xlab("Sales") +
  ylab("Profit")

# -----------------------------------
# Prediction Example
# -----------------------------------

new_data <- data.frame(Sales = c(100, 500, 1000))

predicted_profit <- predict(model, newdata = new_data)

print(predicted_profit)

# -----------------------------------
# Key Insights
# -----------------------------------

cat("Key Insights:\n")
cat("1. Regression model is used to predict profit based on sales.\n")
cat("2. Sales and profit show a relationship in the dataset.\n")
cat("3. Higher sales may lead to higher profit in many cases.\n")
cat("4. Machine learning models help in business decision making.\n")

# -----------------------------------
# Key Skills Used
# -----------------------------------

cat("\nKey Skills Used:\n")
cat("• Basic Machine Learning Concepts\n")
cat("• Regression Modeling\n")
cat("• Model Understanding\n")
cat("• Data Analysis and Visualization\n")
```


